---
title: '2'
author: "SeoYeonYang"
date: "2018년 10월 19일"
output: html_document
---

```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)
library(dplyr)
```

#Lecture 3. Data Transformation.(Ch 5)
dim(flights) : 자료의 행, 열의 개수를 알려줌.<br>
head(flights) : 자료의 상위 6개<br>
tail(flights) : 자료의 하위 6개<br>
colnames(flights) : 자료의 열 이름<br>
<br>
dplyr basics <br>
1. Pick observations by their values: filter(). <br>
2. Reorder the rows: arrange().<br>
3. Pick variables by their names: select().<br>
4. Create new variables with functions of existing variables: mutate().<br>
5. Collapse many values down to a single summary: summarise().<br>

###filter() : Filter with rows
```{r}
jan1 <- filter(flights, month == 1, day == 1) #==가 아닌 =를 쓰면 오류남
filter(flights, month == 11 | month==12 )
filter(flights, month %in% c(11, 12)) #x %in% y : select every row where x is one of the values in y. 위와 동일한 결과

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

#동일한 결과
not_cancelled %>%
  count(dest)
not_cancelled %>%
  group_by(dest) %>%
  summarise(n())

#동일한 결과
not_cancelled %>% count(tailnum, wt = distance) #weighted count
not_cancelled %>%
  group_by(tailnum) %>%
  summarise(n = sum(distance))
```
<br>
<br>

###CF
sqrt(2) ^ 2 == 2 : 컴퓨터의 소수점 한계(근사계산) 때문에 FALSE가 나옴<br>
1 / 49 * 49 == 1 : 얘도 무한 소수라 마찬가지 이유로 FALSE<br>
그래서 near()함수를 쓴다.<br>
<br>
<br>

###NA represents an unknown value
NA > 5		#> [1] NA <br>
10 == NA	#> [1] NA<br>
NA / 2		#> [1] NA<br>
NA == NA	#> [1] NA<br>
If you want to determine if a value is missing, use <is.na()> <br>
<br>
<br>

###arrange() : Arrange rows. 순서를 바꾼다
```{r}
#flights의 자료를 year month day순서를 앞으로 뺀다.
arrange(flights, year, month, day)

#Missing values are always sorted at the end
arrange(flights, desc(dep_delay))		#dep_delay의 내림차순(큰 값부터 작은 값으로)

#배열 조건 여러개도 가능
arrange(flights, desc(is.na(dep_time)), dep_time)
```
<br>
<br>

###select() : Select columns. 데이터 선별
```{r}
select(flights, year, month, day)

#Select all columns except those from year to day (inclusive)
select(flights, -(year:day))

#Select columns by positions, 1열부터 10열까지 2개씩 건너뛰어 고름. (이 경우 홀수 열만 나옴)
select(flights, seq(1, 10, by = 2))

#everything() is useful if you have a handful of variables you’d like to move to the start of the data frame.
select(flights, time_hour, air_time, everything())
```
<br>
<br>

###rename() : Rename a variable
```{r}
rename(flights, tail_num = tailnum) #tailnum을 tail_num으로 바꾼다.
```
<br>
<br>

###Helper functions.
```{r}
#변수명이 “abc”로 시작하는 열 고르기
f1<-flights %>% select(starts_with("abc"))

#변수명이 “xyz”로 끝나는 열 고르기
flights %>% select(ends_with("xyz"))
 
 #변수명에 “ijk” 포함하는 열 고르기
 #contains("ijk")
 
 #selects variables that match a regular expression-특정 패턴에 맞는 정규식 고르기-(more details later).
 #matches("(.)\\1")
 
 #matches x1, x2 and x3.
 #num_range("x", 1:3) 
```
<br>
<br>

###mutate() : Add new variables
```{r}
mutate(flights,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

#If you only want to keep the new variables, use transmute()
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
<br>
<br>

###CF
Modular arithmetic: %/% (integer division) and %% (remainder), where x == y * (x %/% y) + (x %% y) <br>
Logs: log(), log2(), log10()
```{r}
# cumsum(), cumprod(), cummin(), cummax(), cummean().
x<-1:10
cumsum(x) #n번째까지의 합을 출력
cummean(x) #n번째까지의 평균을 출력

# Ranking: start with min_rank() (there are other ranking functions)
#(등수를 출력하는거임. 첫번째 원소 1등, ..., 마지막 원소 5등. NA는 NA)
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)

#이건 제일 큰 값을 가진 변수가 1등
min_rank(desc(y))

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
delays
ggplot(data = delays, mapping = aes(x = tailnum)) + 
  geom_bar()


```

###summarise() : Grouped summaries
```{r}
#mean of a variable
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))

#Useful if paired with group_by():
by_day <- group_by(flights, year, month, day) #자료가 대략 365개의 그룹으로 그루핑됨. But grouping doesn't change how the data looks
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE)) #그룹별로 결과가 나옴. 즉 총 약 365개의 값이 출력됨. 

#count=n()	: count 변수에 group의 멤버수 입력
 delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE) ) %>% 
  filter(count > 20, dest != "HNL") #Unfortunately ggplot2 does not support pipe.
```
<br>

####other summarise functions **벡터 데이터 처리하는 법**
summarise() 함수이 제공하는 수치형 데이터에 대한 요약 통계량 옵션<br>
 - mean(x, na.rm = TRUE) : 평균, 결측값을 제외하고 계산하려면 na.rm = TRUE 추가<br>
 - median(x, na.rm = TRUE) : 중앙값<br>
 - sd(x, na.rm = TRUE) : 표준편차<br>
 - min(x, na.rm = TRUE) : 최소값<br>
 - max(x, na.rm = TRUE) : 최대값<br>
 - IQR(x, na.rm = TRUE) : 사분위수 (Inter Quartile Range = Q3 - Q1)<br>
 - sum(x, na.rm = TRUE) : 합, 결측값을 제외하고 계산하려면 na.rm = TRUE 추가<br>
 - quantile(x, 0.25)<br>
 - first(x), nth(x, 2), last(x)<br>
 - n(x), sum(!is.na(x)), n_distinct(x), count(x)<br>
 - sum(x > 10), mean(y == 0)<br>
 - range(x) x의 범위 출력 <br>
```{r}
#[arr_delay > 0 ] : 참이면 해당하는 열의 숫자를 가져옴
flights%>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay, na.rm=TRUE),
    avg_delay2 = mean(arr_delay[arr_delay > 0], na.rm=TRUE) # the average positive delay
  )


# How many flights left before 5am? (these usually indicate delayed flights from the previous day) ** (dep_time <500) : 참이면 1, 거짓이면 0
flights %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500, na.rm=TRUE))

# What proportion of flights are delayed by more than an hour?
#(arr_delay > 60) : 참이면 1, 거짓이면 0 따라서 비율을 출력하게 됨.
flights %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60, na.rm=TRUE))

# Position: first(x), nth(x, 2), last(x)
flights %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )

# Count: n(x), sum(!is.na(x)), n_distinct(x)
flights %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))

flights %>% 
  count(tailnum, wt = distance) # tailnum에 따라 distance의 합을 구해 출력

#그룹별로 arr_delay가 큰 10개를 출력한다.
flights%>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)

# Find all groups bigger than a threshold:
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 10000 )
dim(popular_dests)
```
<br>
<br>
### 기타
서로 다른 항목의 개수 구하기.
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>% 
  select(year:day, ends_with("dep_time"), ends_with("dep_delay"), dest) %>%
  mutate(., sched_dep_hour = sched_dep_time %/% 100)

popular_flight <- not_cancelled %>% group_by(dest) %>% 
  filter(n() > 10000 )

head(popular_flight)

#주어진 데이터에 dest가 서로 다른게 몇개인가
n_distinct(popular_flight$dest)

#그리고 그것들의 이름은 무엇인가
unique(popular_flight$dest)

popular_flight %>%
  group_by(dest) %>%
  ncol
```

###실습_Grouping by multiple values (Ch 5.6.5)
```{r}
#1
daily <- group_by(flights, year, month, day) #날짜별로 grouping
per_day   <- summarise(daily, flights = n())
per_month <- summarise(per_day, flights = sum(flights)) #summarize : One grouping level will be dropped. 즉 함수를 한 번 적용할 때 마다 맨 마지막으로 group_by에 사용된 그룹 변수를 기준으로 summarize 를 실행한다. 따라서 달별로 붂이게 된다.
per_year  <- summarise(per_month, flights = sum(flights))

#4
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>% 
  select(year:day, ends_with("dep_time"), ends_with("dep_delay"), dest) %>%
  mutate(., sched_dep_hour = sched_dep_time %/% 100)

popular_flight <- not_cancelled %>% group_by(dest) %>% 
  filter(n() > 10000 )

head(popular_flight)

n_distinct(popular_flight$dest)

popular_flight %>%
  group_by(dest) %>%
  ncol#ncol은 grouping에 관계없이 본 데이터의 열개수를 센다.
```




